A Simple Minibatch Acceptance Test for MCMC

### 2016-05-18 nips_2016.tex
Adding another figure for Gaussian Mixtures. Going to sleep tomorrow, will try and do as much writing as possible tomorrow

There is another advantage when using our test for Bayesian
posterior inference. As the dataset size grows, the posterior distribution
sharpens and often concentrates to a few sharp peaks. Samples from
this distribution will come from those peaks, and will easily become
stuck in one of them (see Section~\ref{ssec:gaussians}).  The samples are not exploring
the posterior parameter space, and are unlikely to concentrate at a
strong local optimum. They therefore fail to accurately represent a
sample of the posterior because of inadequate search (i.e.,
mixing). We argue that it is much more natural to sample the
distribution at a higher temperature determined by the
minibatch size rather than the dataset size. This temperature is
determined in a natural way by the variance prerequisite for our
test. Starting at a high temperature (small minibatch size) which
mixes easily, one can reduce the temperature (increasing minibatch
size) until one reaches the desired target distribution temperature.
Although this approach also involves changing minibatch size, the
changes increase minibatch size monotonically, and only a few changes
are performed in the course of a sampling session.
Annealing can also be used for MAP and ML estimation because changing temperature preserves
posterior modes.

@@

In this section, we introduce our new Metropolis-Hastings test and show how to use it to achieve
sufficient mixing and/or to reach better posterior modes. To start, we introduce two key values we
will use frequently, $\Delta$ and $\Delta'$:
\begin{equation}\label{eq:deltas}
\Delta = \log \left( \frac{p(\theta')  \prod_{i=1}^N p(x_i \mid \theta')  q(\theta_t \mid
\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)  q(\theta' \mid
\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta') \prod_{i=1}^n p(x_i \mid \theta') q(\theta_t \mid
\theta')}{p(\theta_t) \prod_{i=1}^n p(x_i \mid \theta_t) q(\theta' \mid
\theta_t)} \right)
\end{equation}
where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}, so the terms here
consider the likelihood ratios in log space. The only difference between $\Delta$ and $\Delta'$ is
that they use $N$ and $n$ terms in the likelihood ratio, respectively. We assume $n \ll N$ (so
$\Delta'$ is substantially faster to compute) and that the $n$ data points $x_i$ represent a random
minibatch of data.

@@

% Daniel: I decided to put the figure here, so it's closer to the text that references it.
\begin{figure}[t]
    \centering
    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
    \caption{
    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
    standard logistic distribution ($\mu=0$, $s=1$); we use three standard deviation (SD) values of
    $X_{\rm norm}$. The red $X_{\rm norm}$ and $X_{\rm corr}$ curves convolve to form the logistic,
    etc. The $y$-axis is capped at 0.5 for readability. This figure must be viewed in color.
    }
    \label{fig:deconvolution}
    \vspace{-10pt}
\end{figure}

@@

\begin{algorithm}[t]
\KwData{$\{x_i\}$ under a target distribution}
\KwResult{A series of sampled $\{\theta_i\}$ under the posterior distribution} 
pre-compute the distribution of $X_{corr}$ under the different value of $var(\Delta')$\;
\While{not enough sampled data}{
	estimate the $sd(\Delta')$\;
	choose the distribution of $X_corr$ from the pre-computed distribution family\;
    propose a candidate $\theta'$\;
    compute the noisy $\Delta'$\;
    sample a $x$ from $X_{corr}$\;
    \eIf{$x + \Delta' > 0$}{
        Accept the $\theta'$;
    }{
        Reject the $\theta'$;
    }
}
\caption{A description of our MH test.}
\label{alg:our_algorithm}
\end{algorithm}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=1\linewidth]{cloud_v01.png}
    \caption{
    {\color{blue}
    Daniel: I'll fix the caption later. Basically, true log posterior, standard/adaptive procedures,
    and our method, using MB=200 out of the 10k samples here. All three used the same proposal
    distribution, a RW with a Gaussian of 0.03.
    }
    }
\end{figure}

@@

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/acc_hyc.png}
        \caption{prediction accuracy}
        \label{fig:logistic_pa}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/ll_hyc.png}
        \caption{loglikelihood}
        \label{fig:logistic_ll}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/roofline_hyc.png}
        \caption{roofline property}
        \label{fig:logistic_roof}
    \end{subfigure}
    \caption{Experiments of logistic regression}\label{fig:logistic}
\end{figure}

@@

Figure~5 illustrates our experiment results.  
\begin{figure}[t]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  %\includegraphics[width=0.5\linewidth]{empty}
  \caption{This will hopefully show results for neural networks on BIDMach.}
\end{figure}

>>>>

There is another advantage when using our test for Bayesian
posterior inference. As the dataset size grows, the posterior distribution
sharpens and often concentrates to a few sharp peaks. Samples from
this distribution will come from those peaks, and will easily become
stuck in one of them (see Section~\ref{ssec:gaussians}).  The samples are not exploring
the posterior parameter %vspace, and are unlikely to concentrate at a
strong local optimum. They therefore fail to accurately represent a
sample of the posterior because of inadequate search (i.e.,
mixing). We argue that it is much more natural to sample the
distribution at a higher temperature determined by the
minibatch size rather than the dataset size. This temperature is
determined in a natural way by the variance prerequisite for our
test. Starting at a high temperature (small minibatch size) which
mixes easily, one can reduce the temperature (increasing minibatch
size) until one reaches the desired target distribution temperature.
Although this approach also involves changing minibatch size, the
changes increase minibatch size monotonically, and only a few changes
are performed in the course of a sampling session.
Annealing can also be used for MAP and ML estimation because changing temperature preserves
posterior modes.

@@

In this section, we introduce our new Metropolis-Hastings test and show how to use it to achieve
sufficient mixing and/or to reach better posterior modes. To start, we introduce two key values we
will use frequently, $\Delta$ and $\Delta'$:
\begin{equation}\label{eq:deltas}
\Delta = \log \left( \frac{p(\theta')  \prod_{i=1}^N p(x_i \mid \theta')  q(\theta_t \mid
\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)  q(\theta' \mid
\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta') \prod_{i=1}^n p(x_i \mid \theta') q(\theta_t \mid
\theta')}{p(\theta_t) \prod_{i=1}^n p(x_i \mid \theta_t) q(\theta' \mid
\theta_t)} \right)
\end{equation}
where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}, so the terms here
consider the likelihood ratios in log %vspace. The only difference between $\Delta$ and $\Delta'$ is
that they use $N$ and $n$ terms in the likelihood ratio, respectively. We assume $n \ll N$ (so
$\Delta'$ is substantially faster to compute) and that the $n$ data points $x_i$ represent a random
minibatch of data.

@@

% Daniel: I decided to put the figure here, so it's closer to the text that references it.
\begin{figure}[t]
    \centering
    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
    \caption{
    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
    standard logistic distribution ($\mu=0$, $s=1$); we use three standard deviation (SD) values of
    $X_{\rm norm}$. The red $X_{\rm norm}$ and $X_{\rm corr}$ curves convolve to form the logistic,
    etc. The $y$-axis is capped at 0.5 for readability. This figure must be viewed in color.
    }
    \label{fig:deconvolution}
    %\vspace{-10pt}
\end{figure}

@@

\begin{algorithm}[t]
\KwData{$\{x_i\}$ under a target distribution}
\KwResult{A series of sampled $\{\theta_i\}$ under the posterior distribution} 
pre-compute the distribution of $X_{corr}$ under the different value of $var(\Delta')$\;
\While{not enough sampled data}{
	estimate the $sd(\Delta')$\;
	choose the distribution of $X_corr$ from the pre-computed distribution family\;
    propose a candidate $\theta'$\;
    compute the noisy $\Delta'$\;
    sample a $x$ from $X_{corr}$\;
    \eIf{$x + \Delta' > 0$}{
        Accept the $\theta'$;
    }{
        Reject the $\theta'$;
    }
}
\caption{A description of our MH test. {\color{blue} TODO}}
\label{alg:our_algorithm}
%\vspace{-10pt}
\end{algorithm}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=1\linewidth]{cloud_v01.png}
    \caption{
    {\color{blue}
    Daniel: I'll fix the caption later. Basically, true log posterior, standard/adaptive procedures,
    and our method, using MB=200 out of the 10k samples here (also 10k passes over data). All three
    used the same proposal distribution, a RW with a Gaussian of 0.03. Temperature is at 110.
    }
    }
    %\vspace{-10pt}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=1\linewidth]{adaptive_and_ours_information_v01.png}
    \caption{
    {\color{blue}
    Daniel: I'll fix the caption later. First plot is capped at 1k but the first bin goes up to
    8000. For the SDs, I skipped the ones that didn't satisfy the requirement (a.k.a. repeating the
    previous sample). The third is when I re-ran the experiments but with different minibatches, to
    see how much I would have to repeat. These values were all with one run, though, I might
    consider just taking several (about five?) and doing error bar stuff.
    }
    }
    %\vspace{-10pt}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/acc_hyc.png}
        \caption{prediction accuracy}
        \label{fig:logistic_pa}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/ll_hyc.png}
        \caption{loglikelihood}
        \label{fig:logistic_ll}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/roofline_hyc.png}
        \caption{roofline property}
        \label{fig:logistic_roof}
    \end{subfigure}
    \caption{Experiments of logistic regression}\label{fig:logistic}
    %\vspace{-10pt}
\end{figure}

@@

Figure~5 illustrates our experiment results.  
\begin{figure}[t]
    \centering
    \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
    %\includegraphics[width=0.5\linewidth]{empty}
    \caption{This will hopefully show results for neural networks on BIDMach.}
    %\vspace{-10pt}
\end{figure}
